Cellular Automata and Computational Universality

### 2018-08-23 paper/paper.tex
Replicator and Diamoeba and paper updates

    Two of these 256 rules are particularly special, namely Rule 54 and 110,
    both of which are Class 4 elementary cellular automata. Rule 110 has been
    proven to be universal, or Turing complete, meaning it is theorectically
    able to simulate any function. Meanwhile, it is so far unknown whether Rule
    54 is capable of universal computation -- interacting structures form, but
    it remains to been seen whether structures useful for computation can be
    formed. Computational universality is discussed more in-depth later on.

@@

    \subsection{Cyclic cellular automata}

\section{Examples}
    \subsection{Von Neumann's automaton}
    \subsection{Game of Life}
    \subsection{Langton's Ant}
    \subsection{WireWorld}
\section{Applications}
\section{Computational Universality}
\section{Firing Squad Synchronisation Problem}
\section{Majority Problem}

>>>>

    Two of these 256 rules are particularly special, namely Rule 54 and 110,
    both of which are Class 4 elementary cellular automata. Rule 110 has been
    proven to be universal, or Turing complete, meaning it is theorectically
    able to simulate any function. Meanwhile, it is so far unknown whether Rule
    54 is capable of universal computation -- interacting structures form, but
    it remains to been seen whether structures useful for computation arise.
    Computational universality is discussed more in-depth later on.

    There have been attempts to classify automata more rigorously, and Culik and
    Yu proposed four well-defined classes for such a task; membership in such
    classes has, however, been shown to be undecidable \cite{CulikYuMembership}.

@@

    Conway originally believed that no pattern could grow infinitely, i.e. for
    the population to grow past any upper limit for some initial configuration
    with finite number of alive cells, and offered fifty dollars to the first
    person who could prove or disprove this conjecture before the end of 1970.
    In November of that year a team from the Massachusetts Institute of
    Technology, led by Bill Gosper, found a structure which produced gliders
    every 30 generations, and called it the Gosper Glider Gun. Later on, smaller
    patterns were found to produce infinite growth, one of which has been proven
    to be minimal, starting off with only ten alive cells
    \cite{infiniteGrowth}. More of what this means for the computational power
    of Game of Life, as well as other automata, will be discussed under
    Computational Universality.

    A cellular automaton is Life-like (similar to Game of Life) if it meets the
    following criteria \cite{LifeLike}:

    \begin{itemize}
        \item it is two-dimensional
        \item each cell of the automaton may be in one of two states, alive or
            dead (or on or off) at any point
        \item a cell's Moore neighbourhood is used to compute the cell's state
            in the next generation
        \item in each time step of the automaton, the next state of the cell can
            be expressed as a funciton of the number of adjacent cells that are
            in the alive state and of the cell's own state
    \end{itemize}

    As with elementary cellular automata, it is helpful to refer to specific
    rulesets in a more generalised way. Where this was using the rule's binary
    encoding for one-dimensional automata, one such convention for
    two-dimensional automat is using a string $Bx/Sy$, where $x$ and $y$ are a
    sequence of digits, 0 to 8, in ascending order. The presence of a digit $d$
    in $x$ means that a dead cell with $d$ alive neighbours will become alive in
    the next generation, while the presence of $d$ in $y$ signifies that an
    alive cell with $d$ alive neighbours will continue to live on to the next
    generation. For example, Game of Life is denoted B3/S23: a dead cell is
    \textbf{B}orn if there are exactly 3 alive neighbours surrounding it, and
    and alive cell \textbf{S}urvives if there are two or three alive cells
    surrounding it.

    \begin{figure}[h]
        \begin{centering}
            \begin{tabular}{cc}
                \includegraphics[width=2in]{replicator1.png} &
                \includegraphics[width=2in]{replicator2.png} \\
                %(a) first & (b) second \\[6pt]
                \includegraphics[width=2in]{replicator3.png} &
                \includegraphics[width=2in]{replicator4.png} \\
                %(c) third & (d) fourth \\[6pt]
                %\multicolumn{2}{c}{\includegraphics[width=65mm]{it} }\\
                %\multicolumn{2}{c}{(e) fifth}
            \end{tabular}
            \caption{Replicator (B1357/S1357) replicating itself over the course of
            56 generations}
        \end{centering}
    \end{figure}

    \subsection{Other types of automata}
    \subsubsection{Reversible cellular automata}
    A cellular automaton is reversible if, for every configuration of the
    automaton, exactly one configuration that causes it (its preimage). If we
    consider the automaton as a function that maps configurations to
    configurations, then for a reversible automaton this function would be
    bijective. Where this is not the case, where not every configuration has a
    preimage, these configurations without preimages are called Garden of Eden
    patterns \cite{GardenOfEden}.
    \subsubsection{Totalistic cellular automata}
    \subsubsection{Proabilistic cellular automata}
    \subsubsection{Cyclic cellular automata}

\section{Computational Universality}

\section{Examples}
    \subsection{Von Neumann's automaton}
    \subsection{Game of Life}
    \subsection{Fractran}
    \subsection{Langton's Ant}
    \subsection{WireWorld}
\section{Applications}
    \subsection{Cryptography}
\section{Firing Squad Synchronisation Problem}
\section{Majority Problem}
